Model Checking Differentially Private Properties

### 2018-01-30 applications.tex
proof adpated
added conclusion
add labels

In pan privacy, data analysis mechanisms are required to respect
privacy even in their internal states~\cite{DNPR:10:DPCO,DNPRY:10:PPSA}.
The idea is to show reachability probabilities are similar on
neighboring observations. In the following, we model a discrete
variant of the mechanisms by Markov decision processes. We will also
specify pan privacy in our model using $\dpCTL$.

@@

Similar to Algorithm~\ref{algorithm:frequency-estimator}, our Markov
decision process has internal states $00$, $01$, $10$, $11$ for the
two bits $f_a$ and $f_b$ of observed event types
(Figure~\ref{figure:adjacent-inputs}). The internal state $01$
represents $(f_a, f_b) = (0, 1)$, for instance. From the
initialization state $I$, the Markov decision process moves to one of
the four internal states randomly with the uniform
distribution. Consider the internal state $01$. If the actions $a[a]$
or $a[b]$ are chosen, they denote that the event type $a$ is currently
observed. The Markov decision process moves to the internal state $01$
and $11$ with probability $.375 = \frac{3}{8}$ and $.625 =
\frac{5}{8}$ respectively. Other transitions are defined similarly.

On the other hand, the initialization state $\underline{I}$
corresponds to the computation on neighbors. The Markov decision
process moves to the interal states $\underline{00}$,
$\underline{01}$, $\underline{10}$, $\underline{11}$ randomly with
the uniform distribution. Consider, for example, the internal state
$\underline{10}$. If the action $a[a]$ or $b[a]$ are chosen, they
denote that the current event type is replaced with the eveny type
$a$. The Markov decision process moves to the internal state
$\underline{00}$ and $\underline{10}$ with probability $.375 =
\frac{3}{8}$ and $.625 = \frac{5}{8}$ respectively. At an internal
state, the Markov decision process can move to the output states $0$,
$1$, or $2$ by the truncated $\frac{1}{2}$-geometric mechanism over
$\{ 0, 1, 2 \}$ (Figure~\ref{figure:randomized-frequency}, \`{a} la
Figure~\ref{figure:geometric-mechanism}). For any action sequence,
observe that paths from the state $I$ and those from $\underline{I}$
formalize the computation of Algorithm~\ref{algorithm:frequency-estimator}
on an observation and its neighbor represented by the sequence. We
hence say the states $I$ and $\underline{I}$ are neighbors in the
Markov decision process (Figure~\ref{figure:frequency-estimator}).

\begin{figure}
  \centering
  \begin{subfigure}{\columnwidth}
    \centering
      \begin{tikzpicture}[->,>=stealth',shorten >=1pt,auto,node
      distance=2cm,node/.style={circle,draw}]
      \node[node] (I) at  ( 0,  0) { $I$ };
      \node[node] (00) at (-1.5,  1.5) { $00$ };
      \node[node] (01) at (-1.5, -1.5) { $01$ };
      \node[node] (10) at ( 1.5,  1.5) { $10$ };
      \node[node] (11) at ( 1.5, -1.5) { $11$ };

@@

      \node[node] (I') at  ( 5.5,    0) { $\underline{I}$ };
      \node[node] (00') at (   4,  1.5) { $\underline{00}$ };
      \node[node] (01') at (   4, -1.5) { $\underline{01}$ };
      \node[node] (10') at (   7,  1.5) { $\underline{10}$ };
      \node[node] (11') at (   7, -1.5) { $\underline{11}$ };

@@

      \node[node] (00) at (-4.5, 1.5) { $00$ };
      \node[node] (01) at (-1.5, 2) { $01$ };
      \node[node] (10) at ( 1.5, 2) { $10$ };
      \node[node] (11) at ( 4.5, 1.5) { $11$ };

      \node[node] (00') at (-4.5, -1.5) { $\underline{00}$ };
      \node[node] (01') at (-1.5, -2) { $\underline{01}$ };
      \node[node] (10') at ( 1.5, -2) { $\underline{10}$ };
      \node[node] (11') at ( 4.5, -1.5) { $\underline{11}$ };

@@

\noindent
\emph{Privacy Specification}.
For classical differential privacy, we are interested in the output
distribution. Algorithm~\ref{algorithm:frequency-estimator} has three
different outputs. By privacy analysis, the state $I$ satisfies the
following $\dpCTLstar$ formula:
\[
\Psi = \dpriv{\ln 4}{0} \F \mathit{at}_0 \wedge \dpriv{\ln 4}{0} \F
\mathit{at}_1 \wedge \dpriv{\ln 4}{0} \F \mathit{at}_2
\]
For pan privacy, we moreover wish to measure privacy loss in internal
states. For observations of length $k$, the probabilities of reaching
each internal states need to be sufficiently close. The state $I$
therefore must satisfy
\[
\Psi
\wedge
\bigwedge_{k \in \bbfN}
\dpriv{\ln 4}{0} \X^k \mathit{at}_{00} \wedge
\dpriv{\ln 4}{0} \X^k \mathit{at}_{01} \wedge
\dpriv{\ln 4}{0} \X^k \mathit{at}_{10} \wedge
\dpriv{\ln 4}{0} \X^k \mathit{at}_{11}.
\]

>>>>

In pan privacy, data analysis mechanisms are required to respect
privacy even in their internal states~\cite{DNPR:10:DPCO,DNPRY:10:PPSA}.
The idea is to show reachability probabilities are similar on
neighboring observations. In the following, we model a discrete
variant of the mechanisms by Markov decision processes. We will also
specify pan privacy in our model using $\dpCTLstar$.

@@

Similar to Algorithm~\ref{algorithm:frequency-estimator}, our Markov
decision process has internal states $s_{00}$, $s_{01}$, $s_{10}$, 
$s_{11}$ for the two bits $f_a$ and $f_b$ of observed event types
(Figure~\ref{figure:adjacent-inputs}). The internal state $s_{01}$
represents $(f_a, f_b) = (0, 1)$, for instance. Moreover, let $xy \in
L (s_{xy})$ for $x, y \in \{ 0, 1 \}$. From the
initialization state $s$, the Markov decision process moves to one of
the four internal states randomly with the uniform
distribution. Consider the internal state $s_{01}$. If the actions $a[a]$
or $a[b]$ are chosen, they denote that the event type $a$ is currently
observed. The Markov decision process moves to the internal state $s_{01}$
and $s_{11}$ with probability $.375 = \frac{3}{8}$ and $.625 =
\frac{5}{8}$ respectively. Other transitions are defined similarly.

On the other hand, the initialization state $t$
corresponds to the computation on neighbors. The Markov decision
process moves to the interal states $t_{00}$,
$t_{01}$, $t_{10}$, $t_{11}$ randomly with
the uniform distribution. Similarly, $xy \in L (t_{xy})$ for $x, y \in
\{ 0, 1 \}$.
Consider, for example, the internal state
$t_{10}$. If the action $a[a]$ or $b[a]$ are chosen, they
denote that the current event type is replaced with the eveny type
$a$. The Markov decision process moves to the internal state
$t_{00}$ and $t_{10}$ with probability $.375 =
\frac{3}{8}$ and $.625 = \frac{5}{8}$ respectively. At an internal
state, the Markov decision process can move to the output states $0$,
$1$, or $2$ by the truncated $\frac{1}{2}$-geometric mechanism over
$\{ 0, 1, 2 \}$ (Figure~\ref{figure:randomized-frequency}, \`{a} la
Figure~\ref{figure:geometric-mechanism}). For any action sequence,
observe that paths from the state $s$ and those from $t$
formalize the computation of Algorithm~\ref{algorithm:frequency-estimator}
on an observation and its neighbor represented by the sequence. We
hence say the states $s$ and $t$ are neighbors in the
Markov decision process (Figure~\ref{figure:frequency-estimator}).

\begin{figure}
  \centering
  \begin{subfigure}{\columnwidth}
    \centering
      \begin{tikzpicture}[->,>=stealth',shorten >=1pt,auto,node
      distance=2cm,node/.style={circle,draw}]
      \node[node] (I) at  ( 0,  0) { $s$ };
      \node[node] (00) at (-1.5,  1.5) { $s_{00}$ };
      \node at (-1.9, 1.9) { $00$ };
      \node[node] (01) at (-1.5, -1.5) { $s_{01}$ };
      \node at (-1.9, -1.9) { $01$ };
      \node[node] (10) at ( 1.5,  1.5) { $s_{10}$ };
      \node at ( 1.9,  1.9) { $10$ };
      \node[node] (11) at ( 1.5, -1.5) { $s_{11}$ };
      \node at ( 1.9, -1.9) { $11$ };

@@

      \node[node] (I') at  ( 5.5,    0) { $t$ };
      \node[node] (00') at (   4,  1.5) { $t_{00}$ };
      \node at ( 3.6,  1.9) { $00$ };
      \node[node] (01') at (   4, -1.5) { $t_{01}$ };
      \node at ( 3.6, -1.9) { $01$ };
      \node[node] (10') at (   7,  1.5) { $t_{10}$ };
      \node at ( 7.4,  1.9) { $10$ };
      \node[node] (11') at (   7, -1.5) { $t_{11}$ };
      \node at ( 7.4, -1.9) { $11$ };

@@

      \node[node] (00) at (-4.5, 1.5) { $s_{00}$ };
      \node at (-4.9,  1.9) { $00$ };
      \node[node] (01) at (-1.5, 2) { $s_{01}$ };
      \node at (-1.9,  2.4) { $01$ };
      \node[node] (10) at ( 1.5, 2) { $s_{10}$ };
      \node at ( 1.9,  2.4) { $10$ };
      \node[node] (11) at ( 4.5, 1.5) { $s_{11}$ };
      \node at ( 4.9,  1.9) { $11$ };

      \node[node] (00') at (-4.5, -1.5) { $t_{00}$ };
      \node at (-4.9, -1.9) { $00$ };
      \node[node] (01') at (-1.5, -2) { $t_{01}$ };
      \node at (-1.9, -2.4) { $01$ };
      \node[node] (10') at ( 1.5, -2) { $t_{10}$ };
      \node at ( 1.9, -2.4) { $10$ };
      \node[node] (11') at ( 4.5, -1.5) { $t_{11}$ };
      \node at ( 4.9, -1.9) { $11$ };

@@

\noindent
\emph{Privacy Specification}.
For classical differential privacy, we are interested in the output
distribution. Algorithm~\ref{algorithm:frequency-estimator} has three
different outputs. By privacy analysis, the state $I$ satisfies the
following $\dpCTLstar$ formula:
\[
\Psi = \dpriv{\ln 4}{0} (\F \mathit{at}_0) \wedge \dpriv{\ln 4}{0} (\F
\mathit{at}_1) \wedge \dpriv{\ln 4}{0} (\F \mathit{at}_2)p
\]
For pan privacy, we moreover wish to measure privacy loss in internal
states. For observations of length $k$, the probabilities of reaching
each internal states need to be sufficiently close. The state $I$
therefore must satisfy
\[
\Psi
\wedge
\bigwedge_{k \in \bbfN}
\dpriv{\ln 4}{0} \X^k {00} \wedge
\dpriv{\ln 4}{0} \X^k {01} \wedge
\dpriv{\ln 4}{0} \X^k {10} \wedge
\dpriv{\ln 4}{0} \X^k {11}.
\]
